Hash Bash: Genetic Algorithms for Hash Function Generation

### 2008-12-04 paper/paper.tex
too damn late

\section{Implementation Details and the Fitness Function}
The genetic algorithm and custom language simulator are implemented completely in Python, with the exception of the Diehard tests \cite{Diehard} which are written in C. There are four primary utility scripts: generate.py, fitness.py, simulate.py, and breed.py. Each has a corresponding parameters file for simple fine-tuning. The script that ties all of these together to compute current and successive generations is master.py, which has its own parameter file. Additionally, we add initial entropy to all functions through the use of a Munroe random number \cite{Munroe}. We initialize certain fixed ranges in memory to these numbers for potential use by a hash function.

\paragraph{}
We make extensive use of Psyco \cite{Psyco} and the processing module \cite{Processing}. Psyco improves individual script execution by utilizing a JIT compiler, while the processing module allows for quick and efficient parallelization of the algorithm itself.
The primary performance bottleneck, the fitness testing, is completely parallelized, with a pool of processes that greedily grab functions to simulate and evaluate. Breeding is less of a performance concern, but it is also parallelized due to the ease with which it is possible. There are numerous additional opportunities for algorithmic parallelization that aren't utilized by our implementation.

@@

\section{Further Thoughts and Conclusion}
Despite all of this, we believe this approach holds promise, and given more time, we feel we can produce more substantial and interesting results. The justification for our optimism goes back to our generator function. As mentioned, it's a trivial hash function that just moves some subset of the input buffer into the output buffer. Indeed, on first glance, it would seem like a poor candidate for an initial function. However, we conducted several runs ($\approx$ 50 generations) with the initial population consisting entirely of this 5-line function. At the end of each of these runs, we noticed that there were two or three long functions ($>$300 lines) whose exact hash outputs varied slightly from their initial forms. The variation wasn't significant, but the fact that phenotypic variation existed at all is a hopeful sign. It stands to reason that if provided additional trivial hash functions that collectively exercised the full power of the assembly language, the generator would help the algorithm to produce substantially more interesting input.

\paragraph{}
Also of note is the fact that the resources available for the purposes of this paper were limited. We only had part-time access to a single dual-core machine that was frequently used for other tasks, and only one (time-strapped) developer.  With the injection of additional human resources, the algorithm could be adapted to run on a cluster, the cloud, etc with the modifications as outlined above implemented. It would definitely allow for more rapid convergence on interesting behavior.
\bibliography{paper_cite}
\end{document}

>>>>

\section{Implementation Details and the Fitness Function}
The genetic algorithm and custom language simulator are implemented completely in Python, with the exception of the Diehard tests \cite{Diehard} which are written in C. There are four primary utility scripts: generate.py, fitness.py, simulate.py, and breed.py. Each has a corresponding parameters file for simple fine-tuning. The script that ties all of these together to compute current and successive generations is master.py, which has its own parameter file. Additionally, we add initial entropy to all hash functions through the use of a Munroe random number \cite{Munroe}. We initialize certain fixed ranges in memory to these numbers for potential use by a hash function.

\paragraph{}
We make extensive use of Psyco \cite{Psyco} and the processing module \cite{Processing}. Psyco improves individual script execution by utilizing a JIT compiler, while the processing module allows for quick and efficient parallelization of the algorithm itself.
The primary performance bottleneck, the fitness testing, is completely parallelized, with a pool of processes that greedily grab functions to simulate and evaluate. Breeding is less of a performance concern, but it is also parallelized due to the ease with which it is possible \cite{Mallya}. There are numerous additional opportunities for algorithmic parallelization that aren't utilized by our implementation.

@@

\section{Further Thoughts and Conclusions}
Despite all of this, we believe this approach holds promise, and given more time, we feel we can produce more substantial and interesting results. The justification for our optimism goes back to our generator function. As mentioned, it's a trivial hash function that just moves some subset of the input buffer into the output buffer. Indeed, on first glance, it would seem like a poor candidate for an initial function. However, we conducted several runs ($\approx$ 50 generations) with the initial population consisting entirely of this 5-line function. At the end of each of these runs, we noticed that there were two or three long functions ($>$300 lines) whose exact hash outputs varied slightly from their initial forms. The variation wasn't significant, but the fact that phenotypic variation existed at all is a hopeful sign. It stands to reason that if provided additional trivial hash functions that collectively exercised the full power of the assembly language, the generator would help the algorithm to produce substantially more interesting input.

\paragraph{}
Also of note is the fact that the resources available for the purposes of this paper were limited. We only had part-time access to a single dual-core machine that was frequently used for other tasks, and only one (very time-strapped) developer.  With the injection of additional human resources, the algorithm could be adapted to run on a cluster, the cloud, etc with the modifications as outlined above implemented. Although the runs conducted with this iteration produced limited results, fixing the deficiencies as outlined above would definitely allow for more rapid convergence on interesting behavior.

\newpage
\bibliography{paper_cite}
\end{document}
